---
Project: "The Nutrition Transition in Ecuador: Processed Food Consumption among Children in a Mixed-indigenous, Rural Community"
Purpose: "Data analysis for determining the effects of ultra-processed food consumption on child growth and development in rural Ecuador" 
Data: "Filename: Lulun2.txt, de-identified data"
Project IRB: "IRB # 201704100" 
Author: "Diane Eapen"
Created: "March 3, 2018"
Latest: "April 25, 2018"
output: html_notebook 
---

```{r}
```

#Packages used for this analysis:

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(foreign)
library(knitr)
library(readr)
library(haven)
library(dplyr)
library(Hmisc)
library(pastecs)

```

#Uploading Lulun II Data

```{r}

LulunIIData_v2 <- read.delim("/Users/dianeelizabetheapen/Desktop/MRF 2016-2018/2017-2018/Ecuador and Sugar Consumption /data/Lulun2.txt")

#Subsetting data to remove cases with missing responses from the group variable
LulunIIData_v2$grupo[LulunIIData_v2$grupo==""] <- NA 

#Converting age from months to years
LulunIIData_v2$ageyears <- LulunIIData_v2$agemos/12

```

#Recoding Data:

##Categorizing BMIz, all according to WHO Growth Standards, 2008

```{r}
attach(LulunIIData_v2)
LulunIIData_v2$bmiz_cat[bmiz_LulunII >1] <- 1 # 1 = "Possible Risk of Overweight", "Overweight" and "Obese"
LulunIIData_v2$bmiz_cat[bmiz_LulunII <=1] <- 0 # 0 = "Normal"
detach(LulunIIData_v2)

#Checking recoded BMIz
table(LulunIIData_v2$bmiz_cat)

#81 = normal 
#50 = those who are either overweight or obese or at possible risk of overweight according to WHO Child Growth Standards
```

##Categorizing HAZ

```{r}

attach(LulunIIData_v2)
LulunIIData_v2$HAZ_cat[haz_LulunII >= -2] <- 0 # 0 = "Normal"
LulunIIData_v2$HAZ_cat[haz_LulunII < -2 ] <- 1 # 1 = "Stunted"
detach(LulunIIData_v2)

#Checking recoded HAZ
table(LulunIIData_v2$HAZ_cat)

67+65
67/132 #Not stunted = 50.8%, in entire sample
65/132 #Stunted = 49.2% 

table(Lulun_Control$HAZ_cat)
30/(30+32) #Not stunted = 48.4%
32/(30+32) #Stunted = 51.6%

stat.desc(Lulun_Control$haz_diff3)
stat.desc(Lulun_Egg$haz_diff3)


table(Lulun_Egg$HAZ_cat)
37/(37+33) #Not stunted = 52.9%
33/(37+33) #Stunted = 47.1%

```

#Subsetting Lulun II Data in to Control and Intervention Groups

```{r}

#Subsetting data into RCT control group and intervention group 
Lulun_Control <- subset(LulunIIData_v2, LulunIIData_v2$grupo == "control")
Lulun_Egg <- subset(LulunIIData_v2, LulunIIData_v2$grupo == "egg")
table
```

##Control Descriptives

```{r}
attach(Lulun_Control)
stat.desc(sex)
stat.desc(weight)
stat.desc(height)
stat.desc(haz_LulunII)
stat.desc(waz_LulunII)
stat.desc(whz_LulunII)
stat.desc(bmiz_LulunII)
stat.desc(bmiz_cat)
detach(Lulun_Control)

attach(Lulun_Control)
stat.desc(haz_baseline)
stat.desc(haz_end)
detach(Lulun_Control)

table(Lulun_Control$sex)
```

##Intervention Descriptives

```{r}
attach(Lulun_Egg)
stat.desc(sex)
stat.desc(weight)
stat.desc(height)
stat.desc(haz_baseline)
stat.desc(haz_end)
stat.desc(haz_LulunII)
stat.desc(waz_LulunII)
stat.desc(whz_LulunII)
stat.desc(bmiz_LulunII)
stat.desc(bmiz_cat)
detach(Lulun_Egg)

table(Lulun_Egg$sex)
range(Lulun_Egg$height)

```

#Packages for logisitc regression

```{r}

library(readr)  #for read txt file
library(knitr)  #for creating nicer tables
library(tidyverse)  # for various packages
library(MASS)  #Ordinal logistic regression
library(erer)  #Average marginal effects for the rdinal logistic regression
```

#Recoded Animal_Own_Tot into a categorical variable (too many animals to one subject, i.e. 500 guinea pigs)

```{r}
stat.desc(LulunIIData_v2$Food_Prod_Tot)
stat.desc(LulunIIData_v2$Animal_Own_Tot)

attach(LulunIIData_v2)
LulunIIData_v2$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 1 & LulunIIData_v2$Animal_Own_Tot <= 5] <- 1 #0-5 animals
LulunIIData_v2$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 5 & LulunIIData_v2$Animal_Own_Tot <= 10] <- 2 #6-10 animals
LulunIIData_v2$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 10 & LulunIIData_v2$Animal_Own_Tot <= 15] <- 3 #10-15 animals
LulunIIData_v2$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 15 & LulunIIData_v2$Animal_Own_Tot <= 20] <- 4 #15-20 animals
LulunIIData_v2$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 20 & LulunIIData_v2$Animal_Own_Tot <= 25] <- 5 #15-20 animals
LulunIIData_v2$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 25] <- 6 #15-20 animals
LulunIIData_v2$animal_own_tot_cat 
detach(LulunIIData_v2)

```

##Test assumptions of linearity, influence, and check percent correctly predicted for BMI in a logistic model

```{r}

# influence plots
# Cook's D plot-identifies observation number in parent dataset
mylogit1<-glm(bmiz_cat ~ juice_sugar, data=LulunIIData_v2, family="binomial") 
  plot(mylogit1, which=4, id.n=5, col="red") 
  
  mylogit2<-glm(HAZ_cat ~ juice_sugar, data=LulunIIData_v2, family="binomial") 
  plot(mylogit2, which=4, id.n=5, col="red") 
  
#checking percent correctly predicted 
xt1<-table(round(predict(mylogit1, type="response")), mylogit1$model$bmiz_cat) 
  xt1  
  
xt2<-table(round(predict(mylogit2, type="response")), mylogit2$model$HAZ_cat) 
  xt2 
  
#Calculating sensitivity and specificity of the model 
xt3<- xt1/sum(xt1)
  xt3
  p<-sum(diag(xt3))
  p #proportion of bmi (yes risk for overweight or overwieght or obese bmi z-score) correctly classified by the model
  
xt4<- xt2/sum(xt2)
  xt4
  p<-sum(diag(xt4))
  p #proportion of haz (yes stunted) correctly classified by the model

```

#Binomial Log Regression Model to explore any unexpected associations

```{r}

#Checking variables in 2x2 tables 
table(LulunIIData_v2$bmiz_cat)
table(LulunIIData_v2$HAZ_cat)
table(LulunIIData_v2$work_you)
table(LulunIIData_v2$relation)
table(LulunIIData_v2$animal_own) #a lot more people who have animals
table(LulunIIData_v2$food_prod) #food prod a lot more people who produce 

#Checking bmiz_cat variable
table(LulunIIData_v2$bmiz_cat)
table(LulunIIData_v2$bmiz_cat)
table(LulunIIData_v2$bmiz_cat)
table(LulunIIData_v2$bmiz_cat)
table(LulunIIData_v2$bmiz_cat)

#Model with sex as a predictor variable to check for potential confounding 
modsex <-glm(bmiz_cat ~ sex, data=LulunIIData_v2, family="binomial") #Box Tidwell technique, test the assumption of linearity
summary(modsex) #Not signficant

```

#Binomial Log Regression Model for the Research Question**

```{r}
#Model with HAZ (a growth indicator and, thus, outcome variable)
modh1 <- glm(HAZ_cat ~ sug_drink + juice_sugar + soda + sug_foods + salty_snacks + animal_own + food_prod + work_you + bmiz_cat, data=LulunIIData_v2, family = "binomial")
summary(modh1)
ORmodh1 <- exp(cbind(OR = coef(modh1), confint(modh1))) 
ORmodh1

#Model with BMIz as outcome variable (a growth indicator as well), and HAZ (for association between the two growth indicators), the 5 junk foods, household food production, animal ownership, and caretakers' employment statuses as the predictor varibales 
#Forced entry regression

modb1 <- glm(bmiz_cat ~ HAZ_cat + sug_drink + juice_sugar + soda + sug_foods + salty_snacks + animal_own + food_prod + work_you, data=LulunIIData_v2, family = "binomial")
summary(modb1) #Note: residual deviance and df are pretty close to each other, therefore, no risk of error from overdispersion


#Odds ratios and 95% CI's
ORmodb1 <- exp(cbind(OR = coef(modb1), confint(modb1))) 
ORmodb1
table(ORmodb1)

#p value and OR 95% CI codes

ORmodb1 <- coef(summary(modb1))

p <- pnorm(abs(table1["t value"]), lower.tail = FALSE) * 2
p

## combined table with OR's and 95% CI's

ORmodb1 <- cbind(table1, `p value` = p)

ORmodb1 <- exp(cbind(OR = coef(modb1), confint(modb1))) 
ORmodb1 

table(ORmodb1)

#Cook's D plot
plot(modb1, which=4, id.n=5, col="red", cex.id=0.60) 

#identify observations with a Cook's D greater than 0.0015 
y<-as.data.frame(cooks.distance(modb1))
colnames(y)[1]<-"CD"
y$obs_no<-rownames(y)
z<-y[which(y$CD>0.0015),]
z$obs_no

#Classification
xt<-table(round(predict(modb1, type="response")), modb1$model$bmiz_cat)
  xt
  xt2<- xt/sum(xt)
  xt2
  p<-sum(diag(xt2))
  p #proportion of bmi (yes risk for overweight or overwieght or obese bmi z-score) correctly classified by the model
  
  #Note: there were 7 influential points in the model 
 
```

#Removal of influential points (exploratory)

```{r}
   
#Removed these observations
LulunIIData_v3 <- read.delim("/Users/dianeelizabetheapen/Desktop/MRF 2016-2018/2017-2018/Ecuador and Sugar Consumption /data/Lulun2_subset1.txt")

#Subsetting data to remove cases with missing responses from the group variable
LulunIIData_v3$grupo[LulunIIData_v2$grupo==""] <- NA 

attach(LulunIIData_v3)
LulunIIData_v3$bmiz_cat[bmiz_LulunII >1] <- 1 # 1 = "Possible Risk of Overweight", "Overweight" and "Obese"
LulunIIData_v3$bmiz_cat[bmiz_LulunII <=1] <- 0 # 0 = "Normal"
detach(LulunIIData_v3)


attach(LulunIIData_v3)
LulunIIData_v3$HAZ_cat[haz_LulunII >= -2] <- 0 # 0 = "Normal"
LulunIIData_v3$HAZ_cat[haz_LulunII < -2 ] <- 1 # 1 = "Stunted"
detach(LulunIIData_v3)


attach(LulunIIData_v3)
LulunIIData_v3$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 1 & LulunIIData_v3$Animal_Own_Tot <= 5] <- 1 #0-5 animals
LulunIIData_v3$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 5 & LulunIIData_v3$Animal_Own_Tot <= 10] <- 2 #6-10 animals
LulunIIData_v3$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 10 & LulunIIData_v3$Animal_Own_Tot <= 15] <- 3 #10-15 animals
LulunIIData_v3$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 15 & LulunIIData_v3$Animal_Own_Tot <= 20] <- 4 #15-20 animals
LulunIIData_v3$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 20 & LulunIIData_v3$Animal_Own_Tot <= 25] <- 5 #15-20 animals
LulunIIData_v3$animal_own_tot_cat[LulunIIData_v2$Animal_Own_Tot > 25] <- 6 #15-20 animals
LulunIIData_v3$animal_own_tot_cat 
detach(LulunIIData_v3)

```

#Same regression from modb1, but without influential points, from subsetted dataset "LulunIIData_v3"

```{r}
modb1_sub <- glm(bmiz_cat ~ HAZ_cat + sug_drink + juice_sugar + soda + sug_foods + salty_snacks + animal_own + food_prod + work_you, data=LulunIIData_v3, family = "binomial")
summary(modb1_sub) #Note: modb1_sub = subsetted data without influential points

#Cook's D plot
plot(modb1_sub, which=4, id.n=5, col="red", cex.id=0.60) 

#identify observations with a Cook's D greater than 0.0015 
y<-as.data.frame(cooks.distance(modb1_sub))
colnames(y)[1]<-"CD"
y$obs_no<-rownames(y)
z<-y[which(y$CD>0.0015),]
z$obs_no

#Note: still not significant assocition between bmiz_cat and UPF's
#Odds ratios and 95% CI's (just to check)
ORmodb1_sub <- exp(cbind(OR = coef(modb1_sub), confint(modb1_sub))) 
ORmodb1_sub
table(ORmodb1_sub)

```

#Stepwise regressions (for exploratory purposes)
##Removal of processed foods from the model to check for potential confounding 

```{r}
modb2 <- glm(bmiz_cat ~ HAZ_cat + food_prod + work_you + relation, data=LulunIIData_v2, family = "binomial")
summary(modb2)
ORmodb2 <- exp(cbind(OR = coef(modb2), confint(modb2))) 
ORmodb2


#Removal of all variables except for caretakers employment status and caretaker's relation to child
modb3 <- glm(bmiz_cat ~ work_you + relation, data=LulunIIData_v2, family = "binomial")
summary(modb3)
ORmodb3 <- exp(cbind(OR = coef(modb3), confint(modb3))) 


#removal of all vairables except for animal ownership
modb4 <- glm(bmiz_cat ~ animal_own, data=LulunIIData_v2, family = "binomial")
summary(modb4)

#removal of all vairables except for household food production
modb5 <- glm(bmiz_cat ~ food_prod, data=LulunIIData_v2, family = "binomial")
summary(modb5)

#model with just household food production and animal owenership
modb6 <- glm(bmiz_cat ~ animal_own + food_prod, data=LulunIIData_v2, family = "binomial")
summary(modb6)

#model with just household food production, HAZ and employment status
modb7 <- glm(bmiz_cat ~ HAZ_cat + food_prod + work_you, data=LulunIIData_v2, family = "binomial")
summary(modb7)

#Same model as modb7 but adding in the junk foods individually
#model with just household food production, HAZ and employment status and sugary drinks
modb8 <- glm(bmiz_cat ~ HAZ_cat + sug_drink + food_prod + work_you, data=LulunIIData_v2, family = "binomial")
summary(modb8)

#model with just household food production, HAZ and employment status and juices with added sugars
modb9 <- glm(bmiz_cat ~ HAZ_cat + juice_sugar + food_prod + work_you, data=LulunIIData_v2, family = "binomial")
summary(modb9)

#model with just household food production, HAZ and employment status and sodas
modb10 <- glm(bmiz_cat ~ HAZ_cat + soda + food_prod + work_you, data=LulunIIData_v2, family = "binomial")
summary(modb10)

#model with just household food production, HAZ and employment status and sugary foods
modb11 <- glm(bmiz_cat ~ HAZ_cat + sug_foods + food_prod + work_you, data=LulunIIData_v2, family = "binomial")
summary(modb11)

#model with just household food production, HAZ and employment status and salty snacks
modb12 <- glm(bmiz_cat ~ HAZ_cat + salty_snacks  + food_prod + work_you, data=LulunIIData_v2, family = "binomial")
summary(modb12)



```
